\documentclass[11pt]{article}
\textwidth 15.9cm
\textheight 23. cm
\addtolength{\oddsidemargin}{-11.5mm}
\addtolength{\evensidemargin}{-11.5mm}
\addtolength{\topmargin}{-19.0mm}

%\usepackage{amsmath}
\usepackage[ansinew]{inputenc}
\usepackage[bbgreekl]{mathbbol}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{url}
\usepackage[numbers,square]{natbib}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{arrows,matrix}
\usepackage{bbm}
\usepackage{authblk}
\usepackage{subcaption}
\renewcommand{\Affilfont}{\footnotesize}
%\renewcommand{\Authfont}{\normalsize}

%\usepackage{algpseudocode}

\usepackage{stmaryrd}
\usepackage{comment}
\usepackage{showlabels}
\usepackage{soul}
%\usepackage[dvipsnames]{xcolor}

\usetikzlibrary{calc,patterns,shapes.geometric}
\usetikzlibrary{arrows,shapes,positioning}
\usetikzlibrary{decorations.markings,decorations.pathmorphing,decorations.pathreplacing}
\usetikzlibrary{decorations.text,decorations.shapes}

%%%%%%%% MACROS
\include{gempic_spectral_macros}

\title{Tests}


\author[1]{The Gempic development team}
\affil[1]{Max-Planck-Institut f\"ur Plasmaphysik, Garching, Germany}


\begin{document}

\maketitle

\begin{abstract}
This note describes some conventions that should be followed in the GEMPIC code
and its documentation.
\end{abstract}

\tableofcontents

There are currently two separate ways to build tests.

The old test style requires input and output files and is dependent on numdiff. It is currently being outphased.

The new test style uses GoogleTest.

\section{Old Test files (Deprecated)}
To write a test code, write a main program in the \verb|Testing| directory corresponding to the code you are testing. This code should be in a file \verb|test_mytest.cpp| and write the results of the test in a file called \verb|test_mytest.output|   
The new test is automatically added to the list of ctests when a \verb|test_mytest.expected_output| file is added in the directory IOFiles1D, IOFiles2D or IOFiles3D in the \verb|Testing| directory containing \verb|test_mytest.cpp|. When ctest is called the files \verb|test_mytest.output| and \verb|test_mytest.expected_output| are compared and the test is passed when they are identical up to round of errors that are set by parameters used in numdiff (these parameters are hard coded in define_ctest.cmake). Which directory of IOFiles1D, IOFiles2D or IOFiles3D is used when ctest is looking for tests is determined at compile time by the value of \verb|AMReX_SPACEDIM|.
Assuming \verb|build| is the directory in which Gempic is built, the executables of the tests are found in \verb|build/Src/MyDir/Testing| and can be run directly without performing all the ctests. You can also run all the tests in a testing directory by executing ctest in build/Src/MyDir/Testing
There are currently two separate ways to build tests.

The old test style requires input and output files and is dependent on numdiff. It is currently being outphased.

The new test style uses GoogleTest.

\section{Old Tests}

\subsection{\verb|test_projection.cpp|}
\subsection{\verb|test_projection_dual.cpp|}
\subsection{\verb|test_projection_polynomial.cpp|}
\subsection{\verb|test_curl_projection.cpp|}
\subsection{\verb|test_hodge_degree2.cpp|}
\subsection{\verb|test_hodge_degree4.cpp|}
\subsection{\verb|test_hodge_degree6.cpp|}
\subsection{\verb|test_plus_minus_mult.cpp|}

\end{document}
