\documentclass[11pt]{article}
\textwidth 15.9cm
\textheight 23. cm
\addtolength{\oddsidemargin}{-11.5mm}
\addtolength{\evensidemargin}{-11.5mm}
\addtolength{\topmargin}{-19.0mm}

\usepackage[ansinew]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{hyperref}

\include{GempicLatexMacros}

\renewcommand{\part}{\texttt{part}}
\newcommand{\cell}{\texttt{cell}}
\newcommand{\node}{\texttt{node}}
\newcommand{\prim}{\texttt{prim}}

\begin{document}

\section{Particle Structure}\label{sec:ParticleStructure}
Particles in GEMPICX are grouped into a vector of shared pointers to species based on the AMReX pure Struct-of-Array (SoA) particle structure (see \href{https://amrex-codes.github.io/amrex/docs_html/Particle.html}{the AMReX particle documentation} for more details). A species is a collection of particles with the same properties and does not have to be a physical species, e.g., multiple electron species with different treatments can coexist in one simulation.
Each species contains the global properties
\begin{itemize}
    \item charge
    \item mass
    \item name
    \item \lstinline{vDim}: The number of velocity dimensions
    \item \lstinline{nData}: The number of additional real data members (typically just weight)
    \item \lstinline{ParticleDataIndices}: Mapping from physical quantities to the data arrays
\end{itemize}
Individual particles then have further properties
\begin{itemize}
    \item \lstinline{id} and \lstinline{cpu}: Constant AMReX-generated identifiers, the combination of which is unique to each particle
    \item positions
    \item velocities
    \item weight
    \item possibly others, depending on the simulation
\end{itemize}
Particles are accessed and manipulated through range-based for loop, \emph{e.g.},
\begin{lstlisting}
for (auto& particleGrid : particleSpecies) {...}
\end{lstlisting}
creates accessors for particle data and distributes work across the different processes available to the simulation. Pointers to particle data are then obtained via this particle iterator. For multi level particle containers a helper struct is used to select individual levels:
\begin{lstlisting}
for (auto& particleGrid : SelectLevel(particleSpecies, lev)) {...}
\end{lstlisting}

In GEMPICX, we use structs of arrays for all particle data. We do not currently provide the ability to have extra integer components.
Since the number of arrays depends on phase space dimension and number of additional variables, the indexing to retrieve a specific quantity is not always the same. \lstinline{ParticleDataIndices} provide a mapping between physical quantities and the corresponding array. Usage is shown in the example below and a full list can be found in \verb|Src/Particle/GEMPIC_Particle.H|.

Since for GPU runs all particle data is stored on the device, the procedure is then first to obtain the pointers, launch a \lstinline{ParallelFor} loop, and access the data inside, \emph{e.g.}
\begin{lstlisting}
// particles is a vector of shared pointers to species
for (auto const& particleSpecies : particles)
{
    for (auto& particleGrid : *particleSpecies)
    {
        // Getting particle number, tile, and index information
        long const np = particleGrid.numParticles();
        auto const ptd = particleGrid.GetParticleTile().getParticleTileData();
        auto const ii = particleSpecies->get_data_indices();

        // other possible prepwork
        ...

        // launch GPU loop over all particles in current tile
        amrex::ParallelFor(np, [=] AMREX_GPU_DEVICE(long pp)
        {
            // do stuff with particles
            ptd.rdata(ii.m_iposx)[pp] = ptd.rdata(ii.m_iweight)[pp] + 0.5 * ptd.rdata(ii.m_ivelx)[pp];
        });
    }
}
\end{lstlisting}


\section{Particle Samplers}
At the moment, particles can be initiated with pseudo-random, quasi-random, and rejection samplers outside of tests.
They are initialized using the \lstinline{Gempic::Particle::init_particles} function.
The different options are selected in the input file (see \ref{sec:ParticleInput}), the details of which follow below.

All the direct methods use random values to
\begin{itemize}
    \item initialize positions from a uniform distribution on the sampling region.
    \item initialize velocities according to one or more Gaussian distributions given in the input file, each specified by mean velocity and the standard deviation of velocity either as a constant or a function in all directions. Initialized by random values through the Box-Muller method.
\end{itemize}
The particle weight is proportional to the mandatory input function in the input file evaluated at the particle position. The particles are spatially uniform and the density profile is represented only in the weights.

The seed of the random generators can be specified in the input file.

\begin{note}
    All samplers can be used on CPU and GPU. \ref{sec:sampler_pseudorandomGPU} is currently the only one running natively and in parallel on device. All others sample on host to a buffer that gets flushed to the device memory. This approach is of coarse less efficient for GPU simulations but allows for easier sampler development that work out of the box on different architectures.
\end{note}

\subsection{Sampler: PseudoRandom}
Generates uniformly distributed random values for position and velocity distributions for all particles cell by cell.

\subsection{Sampler: PseudoRandomBox}
Same as \verb|PseudoRandom| but draws unformly distributed position values on boxes instead of individual cells. There is no guarantee that \verb|nPartPerCell| particles end up in each cell.

\subsection{Sampler: PseudoRandomGPU}
\label{sec:sampler_pseudorandomGPU}
Equivalent to \verb|PseudoRandom| but the sampling is done fully in parallel on device using \verb|amrex::ParallelForRNG|. Also runs in CPU but does not have any advantages in that scenario.

\subsection{Sampler: Sobol}
Positions and velocities are generated based on quasirandom numbers from the Sobol sequence. The sampling is done on the box level and each box uses a random starting index in the sequence determined by the random seed.

\subsection{Sampler: Rejection}
Sampling arbitrary target distribution functions by accepting/rejecting samples of a proposal distribution depending on the ratio of (probability) densities. The proposal distribution is always a Maxwellian that can be configured in the input file and sampled directly. The target distribution is specified by a string \verb|Particle.species.rejectionDistribution| that in addition to the spatial variables \verb|x, y, z| can contain the velocity components \verb|vx, vy, vz|. An example is the standard univariate Maxwellian

\begin{lstlisting}[language=pkgconfig]
    rejectionDistribution = "1 / sqrt(2*3.14159265)^3 * exp(-(vx^2 + vy^2 + vz^2)/(2))"
\end{lstlisting}

but functions can be more involved like a spatial temperature profile at the gyrocenters of particles in a constant magnetic field

\begin{lstlisting}[language=pkgconfig]
    rejectionDistribution = "1 / sqrt(2*3.14159265*(1+0.5*cos(kvarx*(x-vy))))^3 * exp(-(vx^2 + vy^2 + vz^2)/(2*(1+0.5* cos(kvarx*(x-vy)))))"
\end{lstlisting}

To ensure that \textit{target <= proposal}, the proposal is scaled by \verb|Particle.species.rejectionScaling|. To avoid infinite loops, the sampling aborts if \verb|Particle.species.nMaxRejections| samples get rejected in a row. After sampling, some diagnostic information is returned. In addition to the acceptance rate it contains the minimal scaling factor \textit{s} such that \textit{target <= s * proposal}:

\begin{lstlisting}
    Rejection sampling: Acceptance rate                = 0.03997958542
    Rejection sampling: Samples with target > proposal = 0
    Rejection sampling: Minimum scaling necessary      = 23.21383126
\end{lstlisting}

When using parallelization, each process samples globally followed by a particle redistribution.

\begin{note}
    The spatial density of the proposal distribution should be set to \verb|density = 1| to prevent unwanted effects and obtain the expected spatial density of the specified distribution function.

    There are currently the following restrictions on the proposal distribution
    \begin{itemize}
        \item Only a single Gaussian
    \end{itemize}
\end{note}

\subsection{Sampler: RejectionSobol}
Same as \verb|Rejection| but using quasi-random values from the Sobol sequence.

\subsection{Sampler: Juettner}\label{sec:JuettnerSampler}
In the relativistic case, we sample the spatial components of the 4-velocity $\mathbf{u} = \gamma \mathbf{v}$ that follow the Maxwell-Jüttner distribution
$$ f(\mathbf{u}) = \frac{1}{4\pi m^2 c T K_2(mc^2/T)} \exp\left(- \frac{\gamma m c^2}{T}\right)$$
where $\gamma = 1 / \sqrt{1 - \mathbf{v}^2 / c^2}$ is the Lorentz factor, and $K_2(x)$ is the modified Bessel function of the second kind.

The sampling follows the approach from \cite{zenitani2015} (Sobol method + flipping method):
\begin{enumerate}
    \item Sampling the magnitude $|\mathbf{u}|$ by rejection from a gamma proposal distribution
    \item Random scattering on the corresponding surface in momentum space
    \item Applying a Lorentz transformation to shift the distribution
\end{enumerate}

The variables used in the input file are \verb|vThermal|$= \sqrt{T / mc^2}$ and \verb|vMean|$= \beta = v/c$.

\begin{note}
    There are a number of limitations due to the sampling strategy and implementation:
    \begin{itemize}
        \item The temperatures / thermal velocities have to be isotropic
        \item The distribution can only be shifted in a single coordinate direction
        \item The rejecion sampling is very inefficient for weakly relativistic distributions close to a Maxwellian
    \end{itemize}
\end{note}

\subsection{Sampler: JuettnerSobol}
Same as \verb|Juettner| but using quasi-random values from the Sobol sequence.

\end{document}